### 2023-09-18 proofs/main.tex
Finish commenting subtyping

% G ⊨ iN1 ≈au iN2 ⫤ ( Ξ , uM , aus1 , aus2 )
\renewcommand{\ottdruleAUNVarName}[0]{(Var$_{-}^{[[≈au]]}$)}
\renewcommand{\ottdruleAUShiftUName}[0]{($\uparrow^{[[≈au]]}$)}
\renewcommand{\ottdruleAUForallName}[0]{($\forall^{[[≈au]]}$)}
\renewcommand{\ottdruleAUArrowName}[0]{($\rightarrow^{[[≈au]]}$)}
\renewcommand{\ottdruleAUAUName}[0]{(AU$_{-}$)}

@@

\subsection{Equivalence and Normalization}

@@

@@

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
    [>=stealth',node distance=2.6cm,every node/.style={draw,rectangle},every text node part/.style={align=center}]

@@

@@

\begin{algorithm}[Unification]
  \hfill

  \begin{multicols}{2}
  \ottdefnUNUnif{}
  \columnbreak\\
  \ottdefnUPUnif{}
  \end{multicols}
\end{algorithm}

\subsection{Least Upper Bound}
\label{sec:lub}
\ottdefnsLUB

\subsection{Antiunification}
\label{sec:antiunification}
\ottdefnsAU

@@

\subsection{Upper Bounds}
\input{../_build/_gen/tex/lub-lemmas.tex}

\subsection{Upgrade}
\input{../_build/_gen/tex/upgrade-lemmas.tex}

>>>>

% G ⊨ iN1 ≈au iN2 ⫤ ( Ξ , uM , aus1 , aus2 )
\renewcommand{\ottdruleAUNVarName}[0]{(Var$_{-}^{[[≈au]]}$)}
\renewcommand{\ottdruleAUShiftUName}[0]{($\uparrow^{[[≈au]]}$)}
\renewcommand{\ottdruleAUForallName}[0]{($\forall^{[[≈au]]}$)}
\renewcommand{\ottdruleAUArrowName}[0]{($\rightarrow^{[[≈au]]}$)}
\renewcommand{\ottdruleAUAUName}[0]{(AU)}

@@

\subsection{Equivalence and Normalization}
\label{sec:equivalence-normalization}

@@

\Cref{lemma:subt-equiv-algorithmization}
demonstrates that the equivalence of types is the same
as the equality of their normal forms.
\begin{theorempreview}[Correctness of Normalization]
  Assuming the types are well-formed in $[[Γ]]$, 
  \begin{itemize}
    \item [$-$] $[[Γ ⊢ iN ≈ iM]]$ if and only if $[[nf(iN) = nf(iM)]]$;
    \item [$+$] $[[Γ ⊢ iP ≈ iQ]]$ if and only if $[[nf(iP) = nf(iQ)]]$.
  \end{itemize}
\end{theorempreview}

@@

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
    [>={Stealth[scale=2]},node distance=2.6cm,every node/.style={draw,rectangle},every text node part/.style={align=center}]

@@

\begin{itemize}
  \item \ruleref{\ottdruleSCMEPEqEqLabel} and \ruleref{\ottdruleSCMENEqEqLabel}
    are symmetric cases. To merge two matching entries restricting
    a variable to be equivalent to certain types, we check
    that these types are equivalent to each other.
    To do so, it suffices to check for \emph{equality} of their normal forms,
    as discussed in \cref{sec:equivalence-normalization}. 
    After that, we return the left-hand entry.

  \item \ruleref{\ottdruleSCMEEqSupLabel} and \ruleref{\ottdruleSCMESupEqLabel}
    are also symmetric. 
    In this case,
    since one of the entries requires the variable to be equal to 
    a type, the resulting entry must also imply that.
    However, for the soundness, it is needed to ensure that
    the equating restriction is stronger than the subtyping restriction.
    For this purpose, the premise invokes the positive subtyping.

  \item \ruleref{\ottdruleSCMESupSupLabel} 
    In this case, we find the least upper bound of the types from the input
    restrictions, 
    and as the output, restrict the variable to be a supertype of the result.
    The least upper bound procedure will be discussed in \cref{sec:lub}.
\end{itemize}

@@

\begin{algorithm}[Unification]
  \hfill
  \begin{multicols}{2}
  \ottdefnUNUnif{}
  \columnbreak\\
  \ottdefnUPUnif{}
  \end{multicols}
\end{algorithm}

\begin{itemize}
  \item \ruleref{\ottdruleUShiftULabel}, \ruleref{\ottdruleUShiftDLabel}, 
    \ruleref{\ottdruleUForallLabel}, and \ruleref{\ottdruleUExistsLabel}
    are defined congruently. In the shift rules, the algorithm
    removes the outermost constructor. In the
    $[[∀]]$ and $[[∃]]$ rules, it removes the quantifiers,
    adding the quantified variables to the context $[[Γ]]$.
    Notice that $[[Θ]]$, which specifies
    the contexts in which the algorithmic variables must be instantiated,
    is not changed.
  \item \ruleref{\ottdruleUNVarLabel} and \ruleref{\ottdruleUPVarLabel} 
    are the base cases. 
    Since the sides are equal and free from algorithmic variables,
    the unification returns an empty constraint. 
  \item \ruleref{\ottdruleUNVarLabel} and \ruleref{\ottdruleUPVarLabel}
    are symmetric cases constructing the constraints. 
    When an algorithmic variable is unified with a type, 
    we must check that the type is well-formed in the required context,
    and if it is, we return a constraint restricting the variable
    to be equivalent to that type.
  \item \ruleref{\ottdruleUArrowLabel}.
    In this case, the algorithm makes two recursive calls:
    it unifies the arguments and the results of the arrows.
    After that, the resulting constraints are merged using the
    \emph{unification constraint merge} procedure, 
    which is discussed in \cref{sec:constraint-merge}.
    Notice that $[[UC1]]$ and $[[UC2]]$ are guaranteed to be
    \emph{unification} constraints, not arbitrary \emph{subtyping} 
    constraints: it is important for modularizing the proofs, 
    since the properties of the \emph{unification} constraint merge
    can be proved independently from the \emph{subtyping} constraint merge.
\end{itemize}

\subsection{Least Upper Bound}
\label{sec:lub}

In this section, we present
the algorithm finding the least common supertype of two positive types. 
It is used directly by the constraint merge procedure (\cref{sec:constraint-merge}),
and indirectly, through the type upgrade by positive subtyping
(\cref{sec:subtyping}). Perhaps, the least upper bound is the least 
intuitive part of the algorithm, and its correctness will be covered
in \cref{sec:alg-upper-bounds-proofs}.

\begin{algorithm}[The Least Upper Bound Algorithm]
  \hfill\\
  \ottdefnLUBNsub{}
\end{algorithm}

\begin{itemize}
  \item \ruleref{\ottdruleLUBVarLabel}
    The base case is trivial: 
    the least upper bound of to equal variables
    is the variable itself.
  \item \ruleref{\ottdruleLUBShiftLabel}
    In case both sides of the least upper bound are shifted,
    the algorithm needs to find the anti-unifier of them. 
    Intuitively, this is because in general, the upper bounds of
    $[[↓iN]]$ are $[[∃nas.iP]]$ such that 
    $[[nas]]$ can be instantiated with some $[[iMs]]$ so that
    $[[ Γ ⊢ [iMs/nas]iP ≈ ↓iN ]]$ (see \cref{lemma:shape-of-supertypes}).
  \item \ruleref{\ottdruleLUBExistsLabel}
    In this case, we move the quantified variables to the context $[[Γ]]$, 
    and make a recursive call. 
    It is important to make sure that $[[nas]]$ and $[[nbs]]$ are disjoint.
    In this case, it is guaranteed that the resulting 
    $[[fv(iQ)]]$ will be free of $[[nas]]$ and $[[nbs]]$,
    and thus, the resulting type will be a supertype of both sides
    (it will be discussed in \cref{lemma:shape-of-supertypes}).
\end{itemize}

In the positive subtyping algorithm (\cref{sec:subtyping}),
\ruleref{\ottdruleAPUVarLabel} 
generates a restriction of a variable $[[α̂⁺]]$.
On the one hand, this restriction must imply 
$[[â⁺ :≥ iP]]$ for the subtyping to hold.
On the other hand, the type used in this restriction 
must be well-formed in a potentially stronger (smaller) 
context than $[[iP]]$.

To resolve this problem, we define the \emph{upgrade} procedure,
which for given $[[Δ]]$, $[[pnas]]$, and $[[Δ, pnas ⊢ iP]]$,
finds $[[Δ ⊢ iQ]]$---the least supertype of $[[iP]]$ 
among the types well-formed in $[[Δ]]$.

The trick is to make sure that the `forbidden' variables
$[[pnas]]$ are not used explicitly in the supertypes
of $[[iP]]$. For this purpose, we
construct new types $[[iP1]]$ and $[[iP2]]$,
in each of them replacing the forbidden variables
with fresh variables $[[pnbs]]$ and $[[pncs]]$,
and then find the least upper bound of $[[iP1]]$ and $[[iP2]]$.
It turns out that this renaming forces the common types of 
$[[iP1]]$ and $[[iP2]]$ to be agnostic to $[[pnas]]$,
and thus, the supertypes of $[[iP]]$ well-formed in $[[Δ]]$
are exactly the common supertypes of $[[iP1]]$ and $[[iP2]]$.
These properties are considered in more details in \cref{sec:upgrade-lemmas}.

\begin{algorithm}[Type Upgrade]
  \hfill\\
  \ottdefnLUBUp{}
\end{algorithm}

\begin{paragraph}{Note on the Greatest Lower Bound}
  In contrast to the least upper bound, 
  the general greatest lower bound does not exist in our system.
  For instance, consider a positive type $[[iP]]$, 
  together with its non-equivalent
  supertypes $[[iP1]]$ and $[[iP2]] \not\simeq [[iP1]]$
  (for example, $[[iP = ↓↑↓γ⁻]]$, $[[iP1 = ∃α⁻.↓↑↓α⁻]]$, 
  and $[[iP2 = ∃α⁻.↓α⁻]]$).
  Then for arbitrary $[[iQ]]$ and $[[iN]]$, 
  let us consider the common subtypes of 
  $A = [[iQ → ↓↑iQ → ↓↑iQ → iN]]$ and $B = [[iP → ↓↑iP1 → ↓↑iP2 → iN]]$.
  It is easy to see that $[[∀α⁺.∀β⁺. α⁺ → ↓↑α⁺ → ↓↑β⁺ → iN]]$ and 
  $[[∀α⁺.∀β⁺. α⁺ → ↓↑β⁺ → ↓↑α⁺ → iN]]$ are
  both \emph{maximal} common subtypes of $A$ and $B$,
  and since they are not equivalent, none of them is 
  the \emph{greatest} one.

  However, we designed the subtyping system in such a way 
  that the greatest lower bound is not needed:
  the negative variables are always `protected'
  by \emph{invariant} shifts ($[[↑]]$ and $[[↓]]$), 
  and thus, the algorithm can only require
  a substitution of a negative variable to be 
  \emph{equivalent} to some type but never 
  to be a \emph{subtype}.
\end{paragraph}

\subsection{Anti-unification}
\label{sec:antiunification}

Next, we define the anti-unification procedure,
also known as the \emph{most specific generalization}.
As an input, it takes two declarative types
(e.g., in the positive case $[[iP1]]$ and $[[iP2]]$)
and a context $[[Γ]]$.
and returns a type $[[uQ]]$---the generalizer,
containing negative placeholders 
(represented by algorithmic variables) from 
$[[Ξ]]$ and two substitutions $[[uτ1]]$ and $[[uτ2]]$.
The substitutions replace the placeholders with 
declarative types well-formed in $[[Γ]]$,
such that $[[ [uτ1]uQ = iP1 ]]$ and $[[ [uτ2]uQ = iP2 ]]$.
Moreover, the algorithm guarantees that 
$[[uQ]]$ is the most specific type with this property: 
any other generalizer can be turned into $[[uQ]]$ by some substitution
$[[uρ]]$.

It is important to note the differences between 
the standard anti-unification and our version.
First, we only allow the placeholders at \emph{negative} positions,
which means, for example, that $[[α⁺]]$ and $[[β⁺]]$ cannot be
generalized. Second, the generated pair of substitutions 
$[[uτ1]]$ and $[[uτ2]]$ must replace the placeholders with 
types well-formed in a specified context $[[Γ]]$.

The anti-unification algorithm assumes that the input types 
are normalized. This way, anti-unification up-to-equality rather than 
 anti-unification up-to-equivalence is sufficient.

\begin{algorithm}[Anti-unification]
  \hfill

  \ottdefnAUAUP{}
  \ottdefnAUAUN{}

\begin{itemize}
  \item \ruleref{\ottdruleAUPVarLabel} and \ruleref{\ottdruleAUNVarLabel}
    are the base cases. 
    In this case, since the input types are equal, 
    the algorithm returns this type as a generalizer,
    without generating any placeholders.

  \item \ruleref{\ottdruleAUShiftDLabel}, \ruleref{\ottdruleAUShiftULabel},
    \ruleref{\ottdruleAUForallLabel}, and \ruleref{\ottdruleAUExistsLabel}
    are defined congruently. In the shift rules, the algorithm
    removes the outermost constructor. In the
    $[[∀]]$ and $[[∃]]$ rules, it removes the quantifiers.
    Notice that the algorithm does not add the removed variables to
    the context $[[Γ]]$. This is because $[[Γ]]$
    is used to restrict the resulting anti-unification substitutions, 
    and is fixed throughout the algorithm.

  \item \ruleref{\ottdruleAUAULabel} is the most important rule, 
    since it generates the placeholders. 
    This rule only applies if other negative rules failed.
    Because of that, the anti-unification procedure is 
    \emph{not} syntax-directed.

    The generated placeholder is indexed with a pair of 
    types it is mapped to. It allows the algorithm to 
    automatically unite the anti-unification solutions 
    generated by the different branches of 
    \ruleref{\ottdruleAUArrowLabel}.

    Notice that this rule does not have a positive counterpart,
    since we only allow negative placeholders.

  \item \ruleref{\ottdruleAUArrowLabel}
    makes two recursive calls to the anti-unification procedure,
    and unites the results. Suppose that
    $[[uτ1]]$  and $[[uτ2]]$ are the substitutions generated by
    anti-unification of \emph{argument} types of the arrow,
    and $[[uτ1']]$ and $[[uτ2']]$ are the substitutions generated by 
    anti-unification of \emph{result} types of the arrow.
    It is important that if ($[[uτ1]]$, $[[uτ2]]$)
    and ($[[uτ1']]$, $[[uτ2']]$) send some variables to the same pair of types,
    i.e., $[[ [uτ1]α̂⁻ = [uτ1']β̂⁻]]$ and $[[ [uτ2]α̂⁻ = [uτ2']β̂⁻]]$,
    then these variables are equal, i.e., $[[α̂⁻ = β̂⁻]]$.
    This property is guaranteed by \ruleref{\ottdruleAUAULabel}:
    the name of the placeholder is determined by the pair of 
    types it is mapped to.
\end{itemize}

@@

\subsection{Upper Bounds}
\label{sec:alg-upper-bounds-proofs}
\input{../_build/_gen/tex/lub-lemmas.tex}

\subsection{Upgrade}
\label{sec:upgrade-lemmas}
\input{../_build/_gen/tex/upgrade-lemmas.tex}

### 2023-09-18 proofs/tex/au-lemmas.tex
Finish commenting subtyping

\begin{lemma}[Soundness of Anti-Unification] \label{lemma:au-soundness}
    \hfill
    \begin{itemize}
        \item [$+$]  Assuming $[[iP1]]$ and $[[iP2]]$ are normalized,
        if $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$
        then 
        \begin{enumerate}
            \item $[[Γ ; Ξ ⊢ uQ]]$,
            \item $[[Γ ; · ⊢ ausi : Ξ]]$ for $i \in \{1,2\}$
            are anti-unification solutions, and
            \item $[[ [ausi] uQ = iPi ]]$ for $i \in \{1,2\}$.
        \end{enumerate}

        \item [$-$] Assuming $[[iN1]]$ and $[[iN2]]$ are normalized,
        if $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$
        then
        \begin{enumerate}
            \item $[[Γ ; Ξ ⊢ uM]]$,
            \item $[[Γ ; · ⊢ ausi : Ξ]]$ for $i \in \{1,2\}$
            are anti-unification solutions, and
            \item $[[ [ausi] uM = iNi ]]$ for $i \in \{1,2\}$.
        \end{enumerate}
    \end{itemize}
\end{lemma}
\begin{proof}
    We prove it by induction on 
    $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$
    and mutually, $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$.
    Let us consider the last rule applied to infer this judgement.
    \begin{caseof}
        \item \ruleref{\ottdruleAUNVarLabel}, then $[[iN1]] = [[α⁻]] = [[iN2]]$,
              $[[Ξ]] = [[·]]$, $[[uM]] = [[α⁻]]$, and $[[aus1]] = [[aus2]] = [[·]]$.
            \begin{enumerate}
                \item $[[Γ ; · ⊢ α⁻]]$ follows from the assumption $[[Γ ⊢ α⁻]]$,
                \item $[[Γ ; · ⊢ · : ·]]$ holds trivially, and
                \item $[[ [·] α⁻ = α⁻ ]]$ holds trivially.
            \end{enumerate}
        \item \label{case:anti-unification-soundness:shiftu}
         \ruleref{\ottdruleAUShiftULabel}, then $[[iN1]] = [[↑iP1]]$,
                $[[iN2]] = [[↑iP2]]$, and the algorithm makes the recursive call:
                $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$, 
                returning $[[(Ξ, ↑uQ, aus1, aus2)]]$ as the result.

@@

                    \item $[[Γ ; · ⊢ ausi : Ξ]]$ and $[[Γ ; · ⊢ aus'i : Ξ']]$
                        Then $[[Γ ; · ⊢ ausi ∪ aus'i : Ξ ∪ Ξ']]$
                        are well-defined anti-unification solution.
                        Let us take an arbitrary $[[β̂⁻]] \in [[Ξ ∪ Ξ']]$.
                        If $[[β̂⁻]] \in [[Ξ]]$.
                        then $[[Γ ; · ⊢ ausi : Ξ]]$ implies that $[[ausi]]$, and hence,
                        $[[ausi ∪ aus'i]]$ contains an entry well-formed in $[[Γ]]$.
                        If $[[β̂⁻]] \in [[Ξ']]$, the reasoning is symmetric.

                        $[[ausi ∪ aus'i]]$ is a well-defined anti-unification solution:
                        any anti-unification variable occurs 
                        uniquely $[[ausi ∪ aus'i]]$, since by \cref{obs:names-defined-by-mapping},
                        the name of the variable is in one-to-one correspondence with 
                        the pair of types it is mapped to
                        by $[[aus1]]$ and $[[aus2]]$, 
                        an is in one-to-one correspondence with the pair of types it is mapped to
                        by $[[aus'1]]$ and $[[aus'2]]$ 
                        i.e.  if $[[β̂⁻]] \in [[Ξ ∩ Ξ']]$ then $[[ [aus1]β̂⁻ = [aus'1]β̂⁻ ]]$,
                        and $[[ [aus2]β̂⁻ = [aus'2]β̂⁻ ]]$.

@@

\begin{lemma}[Completeness of Anti-Unification] \label{lemma:au-completeness}
    \hfill
    \begin{itemize}
        \item [$+$] 
            Assume that $[[iP1]]$ and $[[iP2]]$ are normalized, and
            there exists $[[(Ξ', uQ', aus'1, aus'2)]]$ such that
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uQ']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$ 
                are anti-unification solutions, and
                \item $[[ [aus'i] uQ' = iPi ]]$ for $i \in \{1,2\}$.
            \end{enumerate}

@@

        \item [$-$] 
            Assume that $[[iN1]]$ and $[[iN2]]$ are normalized, and
            there exists $[[(Ξ', uM', aus'1, aus'2)]]$ such that
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uM']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$,
                are anti-unification solutions, and
                \item $[[ [aus'i] uM' = iNi ]]$ for $i \in \{1,2\}$.
            \end{enumerate}

@@

\begin{lemma}[Initiality of Anti-Unification] \label{lemma:au-initial}
    \hfill
    \begin{itemize}
        \item [$+$] 
            Assume that $[[iP1]]$ and $[[iP2]]$ are normalized, and
            $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$, 
            then $[[(Ξ, uQ, aus1, aus2)]]$ is more specific than
            any other sound anti-unifier $[[(Ξ', uQ', aus'1, aus'2)]]$, i.e.
            if 
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uQ']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$
                are anti-unification solutions , and
                \item $[[ [aus'i] uQ' = iPi ]]$ for $i \in \{1,2\}$
            \end{enumerate}
            then there exists $[[ausr]]$ such that
            $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uQ')]]$ and $[[ [ausr] uQ' = uQ ]]$. 
            Moreover, $[[ [ausr]β̂⁻]]$ can
            be uniquely determined by $[[ [aus'1]β̂⁻ ]]$, $[[ [aus'2]β̂⁻ ]]$, and
            $[[Γ]]$.
        \item [$-$] 
            Assume that $[[iN1]]$ and $[[iN2]]$ are normalized, and
            $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$, 
            then $[[(Ξ, uM, aus1, aus2)]]$ is more specific than
            any other sound anti-unifier $[[(Ξ', uM', aus'1, aus'2)]]$, i.e.
            if
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uM']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$
                are anti-unification solutions, and
                \item $[[ [aus'i] uM' = iNi ]]$ for $i \in \{1,2\}$
            \end{enumerate}
            then there exists $[[ausr]]$ such that
            $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uM')]]$ and $[[ [ausr] uM' = uM ]]$.
            Moreover, $[[ [ausr]β̂⁻]]$ can
            be uniquely determined by $[[ [aus'1]β̂⁻ ]]$, $[[ [aus'2]β̂⁻ ]]$, and
            $[[Γ]]$.
    \end{itemize}
\end{lemma}
\begin{proof}
    First, let us assume that $[[uM']]$ is a algorithmic variable $[[α̂⁻]]$. 
    Then we can take $[[ausr]] = [[α̂⁻]] \mapsto [[uM]]$, which satisfies the required properties:
    \begin{itemize}
        \item $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uM')]]$ holds since 
        $[[Ξ' | uv uM']] = [[{α̂⁻}]]$ and $[[Γ ; Ξ ⊢ uM]]$ by the soundness of anti-unification (\cref{lemma:au-soundness});
        \item $[[ [ausr] uM' = uM ]]$ holds by construction
        \item $[[ [ausr]α̂⁻]] = [[uM]]$ is the anti-unifier of 
            $[[iN1]] = [[ [aus'1] α̂⁻]]$ and $[[iN2]] = [[ [aus'2] α̂⁻]]$
            in context $[[Γ]]$, and hence, it is uniquely determined by them (\cref{obs:au-deterministic}).
    \end{itemize}

>>>>

\begin{lemma}[Soundness of Anti-Unification] \label{lemma:au-soundness}
    \hfill
    \begin{itemize}
        \item [$+$]  Assuming $[[iP1]]$ and $[[iP2]]$ are normalized,
        if $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$
        then 
        \begin{enumerate}
            \item $[[Γ ; Ξ ⊢ uQ]]$,
            \item $[[Γ ; · ⊢ ausi : Ξ]]$ for $i \in \{1,2\}$
            are anti-unification substitutions, and
            \item $[[ [ausi] uQ = iPi ]]$ for $i \in \{1,2\}$.
        \end{enumerate}

        \item [$-$] Assuming $[[iN1]]$ and $[[iN2]]$ are normalized,
        if $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$
        then
        \begin{enumerate}
            \item $[[Γ ; Ξ ⊢ uM]]$,
            \item $[[Γ ; · ⊢ ausi : Ξ]]$ for $i \in \{1,2\}$
            are anti-unification substitutions, and
            \item $[[ [ausi] uM = iNi ]]$ for $i \in \{1,2\}$.
        \end{enumerate}
    \end{itemize}
\end{lemma}
\begin{proof}
    We prove it by induction on 
    $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$
    and mutually, $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$.
    Let us consider the last rule applied to infer this judgement.
    \begin{caseof}
        \item \ruleref{\ottdruleAUNVarLabel}, then $[[iN1]] = [[α⁻]] = [[iN2]]$,
              $[[Ξ]] = [[·]]$, $[[uM]] = [[α⁻]]$, and $[[aus1]] = [[aus2]] = [[·]]$.
            \begin{enumerate}
                \item $[[Γ ; · ⊢ α⁻]]$ follows from the assumption $[[Γ ⊢ α⁻]]$,
                \item $[[Γ ; · ⊢ · : ·]]$ holds trivially, and
                \item $[[ [·] α⁻ = α⁻ ]]$ holds trivially.
            \end{enumerate}
        \item \label{case:anti-unification-soundness:shiftu}
         \ruleref{\ottdruleAUShiftULabel}, then $[[iN1]] = [[↑iP1]]$,
                $[[iN2]] = [[↑iP2]]$, and the algorithm makes the recursive call:
                $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$, 
                returning $[[(Ξ, ↑uQ, aus1, aus2)]]$ as the result.

@@

                    \item $[[Γ ; · ⊢ ausi : Ξ]]$ and $[[Γ ; · ⊢ aus'i : Ξ']]$
                        Then $[[Γ ; · ⊢ ausi ∪ aus'i : Ξ ∪ Ξ']]$
                        are well-defined anti-unification substitutions.
                        Let us take an arbitrary $[[β̂⁻]] \in [[Ξ ∪ Ξ']]$.
                        If $[[β̂⁻]] \in [[Ξ]]$.
                        then $[[Γ ; · ⊢ ausi : Ξ]]$ implies that $[[ausi]]$, and hence,
                        $[[ausi ∪ aus'i]]$ contains an entry well-formed in $[[Γ]]$.
                        If $[[β̂⁻]] \in [[Ξ']]$, the reasoning is symmetric.

                        $[[ausi ∪ aus'i]]$ is a well-defined anti-unification substitution:
                        any anti-unification variable occurs 
                        uniquely $[[ausi ∪ aus'i]]$, since by \cref{obs:names-defined-by-mapping},
                        the name of the variable is in one-to-one correspondence with 
                        the pair of types it is mapped to
                        by $[[aus1]]$ and $[[aus2]]$, 
                        an is in one-to-one correspondence with the pair of types it is mapped to
                        by $[[aus'1]]$ and $[[aus'2]]$ 
                        i.e.  if $[[β̂⁻]] \in [[Ξ ∩ Ξ']]$ then $[[ [aus1]β̂⁻ = [aus'1]β̂⁻ ]]$,
                        and $[[ [aus2]β̂⁻ = [aus'2]β̂⁻ ]]$.

@@

\begin{lemma}[Completeness of Anti-Unification] \label{lemma:au-completeness}
    \hfill
    \begin{itemize}
        \item [$+$] 
            Assume that $[[iP1]]$ and $[[iP2]]$ are normalized, and
            there exists $[[(Ξ', uQ', aus'1, aus'2)]]$ such that
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uQ']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$ 
                are anti-unification substitutions, and
                \item $[[ [aus'i] uQ' = iPi ]]$ for $i \in \{1,2\}$.
            \end{enumerate}

@@

        \item [$-$] 
            Assume that $[[iN1]]$ and $[[iN2]]$ are normalized, and
            there exists $[[(Ξ', uM', aus'1, aus'2)]]$ such that
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uM']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$,
                are anti-unification substitutions, and
                \item $[[ [aus'i] uM' = iNi ]]$ for $i \in \{1,2\}$.
            \end{enumerate}

@@

\begin{lemma}[Initiality of Anti-Unification] \label{lemma:au-initial}
    \hfill
    \begin{itemize}
        \item [$+$] 
            Assume that $[[iP1]]$ and $[[iP2]]$ are normalized, and
            $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$, 
            then $[[(Ξ, uQ, aus1, aus2)]]$ is more specific than
            any other sound anti-unifier $[[(Ξ', uQ', aus'1, aus'2)]]$, i.e.
            if 
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uQ']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$, and
                \item $[[ [aus'i] uQ' = iPi ]]$ for $i \in \{1,2\}$
            \end{enumerate}
            then there exists $[[ausr]]$ such that
            $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uQ')]]$ and $[[ [ausr] uQ' = uQ ]]$. 
            Moreover, $[[ [ausr]β̂⁻]]$ can
            be uniquely determined by $[[ [aus'1]β̂⁻ ]]$, $[[ [aus'2]β̂⁻ ]]$, and
            $[[Γ]]$.
        \item [$-$] 
            Assume that $[[iN1]]$ and $[[iN2]]$ are normalized, and
            $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$, 
            then $[[(Ξ, uM, aus1, aus2)]]$ is more specific than
            any other sound anti-unifier $[[(Ξ', uM', aus'1, aus'2)]]$, i.e.
            if
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uM']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$, and
                \item $[[ [aus'i] uM' = iNi ]]$ for $i \in \{1,2\}$
            \end{enumerate}
            then there exists $[[ausr]]$ such that
            $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uM')]]$ and $[[ [ausr] uM' = uM ]]$.
            Moreover, $[[ [ausr]β̂⁻]]$ can
            be uniquely determined by $[[ [aus'1]β̂⁻ ]]$, $[[ [aus'2]β̂⁻ ]]$, and
            $[[Γ]]$.
    \end{itemize}
\end{lemma}
\begin{proof}
    First, let us assume that $[[uM']]$ is a algorithmic variable $[[α̂⁻]]$. 
    Then we can take $[[ausr]] = [[α̂⁻]] \mapsto [[uM]]$, which satisfies the required properties:
    \begin{itemize}
        \item $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uM')]]$ holds since 
        $[[Ξ' | uv uM']] = [[{α̂⁻}]]$ and $[[Γ ; Ξ ⊢ uM]]$ by the soundness of anti-unification (\cref{lemma:au-soundness});
        \item $[[ [ausr] uM' = uM ]]$ holds by construction
        \item $[[ [ausr]α̂⁻]] = [[uM]]$ is the anti-unifier of 
            $[[iN1]] = [[ [aus'1] α̂⁻]]$ and $[[iN2]] = [[ [aus'2] α̂⁻]]$
            in context $[[Γ]]$, and hence, it is uniquely determined by them (\cref{obs:au-deterministic}).
    \end{itemize}
